---
title: "Estimating the posterior with a binomial likelihood"
author:
  - "Jeffrey Arnold"
  - "Connor Gilroy"
date: "2018-04-10"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
set.seed(123)
```

# Data

A series of $n$ trials with $y$ successes. We're estimating the probability of success, $\theta$.

```{r}
n <- 10
y <- 5
```

# Grid approximation

Using a uniform prior. Adapted from Richard McElreath, Rethinking Statistics.

```{r}
grid_size <- 100
p_grid <- seq(from = 0, to = 1, length.out = grid_size)
prior <- rep(1, grid_size)
likelihood <- dbinom(x = y, size = n, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)

plot(p_grid, posterior, type = "b")
```

# Analytic solution

Writing the prior as a beta distribution, because the beta distribution is the conjugate prior distribution for the binomial likelihood. 

If $\alpha = \beta = 1$, the beta distribution is uniform between 0 and 1.

```{r}
a <- 1
b <- 1

apost <- y + 1
bpost <- n - y + 1

plot(p_grid, dbeta(p_grid, apost, bpost))
```

```{r}
nsims <- 10000
posterior_samples <- rbeta(nsims, apost, bpost)
mean(posterior_samples)
median(posterior_samples)
```

# Sampling using Stan

```{r}
library("rstan")

options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

```{r}
binomial_model <- stan_model("binomial.stan")
```

```{r}
binomial_model
```

```{r}
binomial_fit <- sampling(binomial_model, data = list(y = y, n = n))
```

You can combine `stan_model()` and `sampling()` into one step by just running `stan()`. 

```{r}
binomial_fit
```

```{r}
plot(binomial_fit)
```

For more plotting methods, we can use the `bayesplot` package, which we will discuss in more detail later.

```{r}
library("bayesplot")
mcmc_dens(as.array(binomial_fit), pars = "theta") + xlim(0, 1)
```
